Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-26 article.tex
Template options
minor things

 % The following command is just for this sample document:
\newcommand{\cs}[1]{\texttt{\char`\\#1}}

 % change the arguments, as appropriate, in the following:
%\jmlrvolume{1}
%\jmlryear{2014}
%\jmlrworkshop{\textcolor{red}{Neural Connectomics Workshop}}
\jmlrproceedings{}{}

@@

 % Authors with different addresses:
 % \author{\Name{Author Name1} \Email{abc@sample.com}\\
 % \addr Address 1
 % \AND
 % \Name{Author Name2} \Email{xyz@sample.com}\\
 % \addr Address 2
 %}

\editor{Editor's name}
 % \editors{List of editors' names}

@@

Given this assumption, we then propose to use as a measure of the
strength of the connection between two neurons $i$ and $j$, the
partial correlation coefficient $p_{i,j}$ between their corresponding
random variables $X_i$ and $X_j$, defined by \cite{}:
\begin{equation}
p_{i,j} =
-\frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}}, \label{eq:inverse}
\end{equation}
where $\Sigma^{-1}$, known as the precision or concentration matrix, is the inverse of the covariance matrix $\Sigma$ of $X$.%%  Partial
%% correlation can be interpreted in several ways. The partial
%% correlation coefficient $p_{i,j}$ can be shown to measure the
%% correlation between $X_i$ and $X_j$ when they are corrected for the
%% (linear) effect of all other variables in $X$ \cite{}.
Assuming that the distribution $P_X$ is a multivariate Gaussian
distribution ${\cal N}(\mu,\Sigma)$, it can be shown that $p_{i,j}$ is
zero if and only if $X_i$ and $X_j$ are independent given all other
variables in $X$, ie. $X_i \perp X_j|X^{-i,j}$ where $X^{-i,j}= X
\setminus\{X_i,X_j\}$. Partial correlation thus measures conditional
dependencies between variables. As an association measures between
neurons, it should thus naturally only detect direct associations
between neurons and filter out spurious indirect effects. The interest
of partial correlation as an association measure has already been
shown for the inference of gene regulatory networks \cite{xxx}.
%% Given a set of multivariate Gaussian
%% random variables $X$, the (undirected) graph that connects all pairs
%% of variables $X_i$ and $X_j$ such that $p_{i,j}\neq 0$ is called a
%% Gaussian graphical models (GGMs). Because it measures conditional
%% dependencies, partial correlation should naturally only detect direct
%% associations between neurons. The superiority of partial correlation
%% over standard correlation as an association measure have already been
%% shown in the context of the inference of gene regulatory networks from
%% gene expression data \cite{}. We will compare both measures in Section
%% \ref{sec:results}.
Note that the partial correlation statistic is symmetric
(i.e. $p_{i,j}=p_{j,i}$). Therefore, our approach cannot identify the
direction of the interactions between neurons. We will see in
Section~\ref{sec:results} why this only affects slightly its
performances, with respect to the metric used in the Connectomics
challenge.

@@

\paragraph{Evaluation of the method}

>>>>

\jmlryear{2014}
\jmlrworkshop{Neural Connectomics Workshop: From Imaging to Connectivity}
\jmlrauthors{A. Sutera et al}

@@

@@

Given this assumption, we then propose to use as a measure of the
strength of the connection between two neurons $i$ and $j$, the
partial correlation coefficient $p_{i,j}$ between their corresponding
random variables $X_i$ and $X_j$, defined by \cite{}:
\begin{equation}
p_{i,j} =
-\frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}}, \label{eq:inverse}
\end{equation}
where $\Sigma^{-1}$, known as the precision or concentration matrix, is the inverse of the covariance matrix $\Sigma$ of $X$. %%  Partial
%% correlation can be interpreted in several ways. The partial
%% correlation coefficient $p_{i,j}$ can be shown to measure the
%% correlation between $X_i$ and $X_j$ when they are corrected for the
%% (linear) effect of all other variables in $X$ \cite{}.
Assuming that the distribution $P_X$ is a multivariate Gaussian
distribution ${\cal N}(\mu,\Sigma)$, it can be shown that $p_{i,j}$ is
zero if and only if $X_i$ and $X_j$ are independent given all other
variables in $X$, ie. $X_i \perp X_j|X^{-i,j}$ where $X^{-i,j}= X
\setminus\{X_i,X_j\}$. Partial correlation thus measures conditional
dependencies between variables. As an association measures between
neurons, it should thus naturally only detect direct associations
between neurons and filter out spurious indirect effects. The interest
of partial correlation as an association measure has already been
shown for the inference of gene regulatory networks \cite{xxx}.
%% Given a set of multivariate Gaussian
%% random variables $X$, the (undirected) graph that connects all pairs
%% of variables $X_i$ and $X_j$ such that $p_{i,j}\neq 0$ is called a
%% Gaussian graphical models (GGMs). Because it measures conditional
%% dependencies, partial correlation should naturally only detect direct
%% associations between neurons. The superiority of partial correlation
%% over standard correlation as an association measure have already been
%% shown in the context of the inference of gene regulatory networks from
%% gene expression data \cite{}. We will compare both measures in Section
%% \ref{sec:results}.
Note that the partial correlation statistic is symmetric
(i.e. $p_{i,j}=p_{j,i}$). Therefore, our approach cannot identify the
direction of the interactions between neurons. We will see in
Section~\ref{sec:results} why this only affects slightly its
performances, with respect to the metric used in the Connectomics
challenge.

@@

\paragraph{Evaluation of the method.}
